Improving content discovery through combining linked data and data mining techniques

### 2015-11-11 report/report.tex
Starting defining the data pipeline.

@@

\begin{lstlisting}[label=lst:sparqlfoaf,caption={Example SPARQL query for people's names and email addresses},frame=single,captionpos=b,language=sparql]
PREFIX foaf: <http://xmlns.com/foaf/0.1/>
SELECT ?name ?email
WHERE {
  ?person a foaf:Person.
  ?person foaf:name ?name.
  ?person foaf:mbox ?email.
}
\end{lstlisting}

@@

\begin{lstlisting}[label=lst:sparqlabout,caption={SPARQL query to determine what },frame=single,captionpos=b,language=sparql]
PREFIX rdf: <http://www.w3.org/1999/02/22-rdf-syntax-ns#>
SELECT ?topic
WHERE {
  ?article rdf:about ?topic .
}
\end{lstlisting}

@@

\section{RDF in the enterprise}

@@

\section{Dataflow Model}

@@

        URI -> RDF [label="extract"];
        RDF -> Features [label="generate"];
      }
    \end{dot2tex}
  \end{center}
  \caption{Null content miner}
\end{figure}

@@

        URI -> RDFa [label="fetch"];
        RDFa -> RDF [label="parse"];
        RDF -> Features [label="generate"];
      }
    \end{dot2tex}
  \end{center}
  \caption{Semantic web content miner}
\end{figure}

@@

        URI [label="URI"];
        RDF1 [label="RDF_1"];
        RDF2 [label="RDF_2"];
        RDFp [label="RDF'"];

        URI -> RDFa [label="fetch"];
        RDFa -> RDF [label="parse"];
        RDF -> RDF1 [label="DBPedia"];
        RDF -> RDF2 [label="DBPedia"];
        RDF -> RDFp [label="\cup"];
        RDF1 -> RDFp [label="\cup"];
        RDF2 -> RDFp [label="\cup"];
        RDFp -> Features [label="generate"];
      }
    \end{dot2tex}
  \end{center}
  \caption{Semantic web content miner with DBPedia enrichment}
\end{sidewaysfigure}

@@

        URI -> RDFa [label="parse"];
        URI -> Entities [label="extract"];
        URI -> Linked [label="scrape"];

@@

\chapter{Implementation}
\chapter{Evaluation}
\chapter{Conclusion}

>>>>

\colorlet{punct}{red!60!black}
\definecolor{background}{HTML}{EEEEEE}
\definecolor{delim}{RGB}{20,105,176}
\colorlet{numb}{magenta!60!black}

\lstdefinelanguage{json}{
    basicstyle=\normalfont\ttfamily\footnotesize,
    numbersep=8pt,
    showstringspaces=false,
    breaklines=true,
    literate=
     *{0}{{{\color{numb}0}}}{1}
      {1}{{{\color{numb}1}}}{1}
      {2}{{{\color{numb}2}}}{1}
      {3}{{{\color{numb}3}}}{1}
      {4}{{{\color{numb}4}}}{1}
      {5}{{{\color{numb}5}}}{1}
      {6}{{{\color{numb}6}}}{1}
      {7}{{{\color{numb}7}}}{1}
      {8}{{{\color{numb}8}}}{1}
      {9}{{{\color{numb}9}}}{1}
      {:}{{{\color{punct}{:}}}}{1}
      {,}{{{\color{punct}{,}}}}{1}
      {\{}{{{\color{delim}{\{}}}}{1}
      {\}}{{{\color{delim}{\}}}}}{1}
      {[}{{{\color{delim}{[}}}}{1}
      {]}{{{\color{delim}{]}}}}{1},
}

\lstset{frame=single,captionpos=b}

@@

\begin{lstlisting}[label=lst:sparqlfoaf,caption={Example SPARQL query for people's names and email addresses},language=sparql]
PREFIX foaf: <http://xmlns.com/foaf/0.1/>
SELECT ?name ?email
WHERE {
  ?person a foaf:Person.
  ?person foaf:name ?name.
  ?person foaf:mbox ?email.
}
\end{lstlisting}

@@

\begin{lstlisting}[label=lst:sparqlabout,caption={SPARQL query to determine what },language=sparql]
PREFIX rdf: <http://www.w3.org/1999/02/22-rdf-syntax-ns#>
SELECT ?topic
WHERE {
  ?article rdf:about ?topic .
}
\end{lstlisting}

@@

\section{RDF in the enterprise}
\label{sec:linked-enterprise-data}

@@

\section{Use Cases}

\section{Data Pipeline}

A core subsystem in the overall system is a conceptual data pipeline whose
input is a URI or IRI identifying a content item published on an organisation's
website and the output is feature sets ready for applying machine learning.

In this section, a theoretical pipeline is inductively defined in steps such
that an application of this pipeline would choose to implement some subset of
all potential pipeline stages as appropriate for the relevant problem domain.

In Chapter~\ref{chp:implementation}, aa system is engineered that implements
as many of these pipeline stages as possible such that a running instance of
the application can configure which components to use and which not to use.
Then in Chapter~\ref{chp:evaluation}, an evalution of the system is given
while it is running each component in isolation to demonstrate which of the
theoretically-defined processes in this chapter appears to be most effective
in generating feature sets specifically for clustering web content.

This system requires some initial definition of some data structures in use:

\begin{description}

\item[IRI] \hfill \\
The input to the system is a character string conformant to the
IRI syntax defined in RFC 3987\footnote{http://tools.ietf.org/html/rfc3987}.
This allows more generality offered by
URIs\footnote{http://tools.ietf.org/html/rfc3986} but is trivially made
compatible with systems that use URIs through the conversion algorithm defined
in section~3.2 of RFC 3987. Note that the public URL by which the public can
read or otherwise consume the content is a valid identifier, but we are not
restricted to that.

\item[Feature Set] \hfill \\
The final output of this data pipeline is a data structure
analogous to a relation or tuple per IRI fed into the system. Every IRI should
have a literal value against all possible columns or fields. For binary fields,
(e.g. the presence of absence of a concept tag), a more pragmatic structure
might be a list of tags positively associated with the IRI rather than
explicitly assigning $false$ to all tags to which the content item does not
pertain. This is analogous to a spare matrix when dealing with a large number
of dimensions.

\item[Named RDF Graph] \hfill \\
The structure used throughout most of the data pipeline
is that of an RDF graph. This is used for all the benefits outlined in
Section~\ref{sec:linked-enterprise-data} such as ease of transformation and
combining of data sets. Named graphs are used such that all data acquired
are keyed back to the IRI of the content item being processed. This also allows
all graphs to be combined in a \emph{triplestore} if needed to allow SPARQL
queries across the combined data for all content items. This can be modelled as
a data structure in many programming languages, but where a serialisation is
used (e.g. examples shown here or to send the data between components), the
JSON-LD\cite{sporny2014json} syntax will be used.
\end{description}

\begin{centering}
\begin{lstlisting}[label=lst:jsonld-identity,caption={Identity graph for a content item in JSON-LD syntax},language=json]
{
  "@id": "http://example.com/entity/1",
  "@graph": []
}
\end{lstlisting}
\end{centering}

With the knowledge only of a content item's IRI, we are arguably only able to
produce an empty named RDF graph. Such a graph for an example IRI
\texttt{http://example.com/entity/} is illustrated in
JSON-LD syntax in Listings~\ref{lst:jsonld-identity}.

The most na\"ive feature set we can generate from such an RDF graph is clearly
a singleton relation \texttt{("http://example.com/entity/")} where a single
$IRI$ field has the value \texttt{"http://example.com/entity/"}. It is also
clear that a set of one-dimension feature vectors with unique values in each
is not suitable for any form of machine learning activity. This does, however,
illustrate a baseline for a working software application that is -- at least
in the syntactic sense -- transforming IRI inputs to feature sets outputs.
Such a $null$ feature generator is depicted in Figure~\ref{fig:gen-null}.

@@

        IRI -> RDF [label="extract"];
        RDF -> Features [label="generate"];
      }
    \end{dot2tex}
  \end{center}
  \caption{Null feature generator \label{fig:gen-null}}
\end{figure}

Note that Figure~\ref{fig:gen-null} shows all three data structures involved
despite having no functional use. We can also see top-level definitions of the
process where we first \emph{extract} semantic information in RDF from a
content item indentified by IRI and then \emph{generate} features therefrom.
More useful models can now be inductively defined by adding atomic
subcomponents that may each add value to the overall transformation.

There are two clear axes along which we can improve this pipeline: expand
the size of the RDF graph we are extracting and improve how we convert this
graph into features. In the first instance, we can consider the former and add
a single pipeline stage for expanding the RDF graph.

Tim Berners-Lee outlined four rules\cite{berners2011linked} for Linked Data,
rule number three of which states ``When someone looks up a URI, provide useful
information, using the standards''. If we assume that many pages have embedded
some semantic web or RDF data, then a simple extraction strategy would be
to deference the content item's IRI via an HTTP GET and pass the content
to a parser capable of extracting RDFa, microformats, etc.

@@

        IRI -> RDFa [label="fetch"];
        RDFa -> RDF [label="parse"];
        RDF -> Features [label="generate"];
      }
    \end{dot2tex}
  \end{center}
  \caption{Semantic web content miner}
\end{figure}

@@

        IRI [label="IRI"];
        RDF1 [label="RDF_1"];
        RDF2 [label="RDF_2"];
        RDFp [label="RDF'"];

        IRI -> RDFa [label="fetch"];
        RDFa -> RDF [label="parse"];
        RDF -> RDF1 [label="DBPedia"];
        RDF -> RDF2 [label="DBPedia"];
        RDF -> RDFp [label="\cup"];
        RDF1 -> RDFp [label="\cup"];
        RDF2 -> RDFp [label="\cup"];
        RDFp -> Features [label="generate"];
      }
    \end{dot2tex}
  \end{center}
  \caption{Semantic web content miner with DBPedia enrichment}
\end{sidewaysfigure}

@@

        IRI -> RDFa [label="parse"];
        IRI -> Entities [label="extract"];
        IRI -> Linked [label="scrape"];

@@

\chapter{Implementation}
\label{chp:implementation}

\chapter{Evaluation}
\label{chp:evaluation}

\chapter{Conclusion}
